\section{Goals}
\label{text:introduction/goals}
In this work, I want to tackle safe navigation in human crowds by combining two of the main paradigms in this field, trajectory optimization algorithms, and deep learning prediction models. I want to come up with an algorithm that both takes the complex human behavior into account, without the need of strong assumptions about human behavior, while still being able to track and guarantee properties of the resulting robot trajectory. Also, I want to re-frame the problem of safe navigation from pure travel-time or safety optimally to an interaction-aware objective. Instead of the control effort, the disturbance the robot causes on the human behavior should be minimized, e.g., not to disrupt doctors in a hospital floor or the pedestrians strolling in the park mentioned above, while guaranteeing safety and reaching the goal when possible.
\newline
Therefore I will use a state-of-the-art deep learned probabilistic and multi-modal pedestrian trajectory prediction model combined with a shooting trajectory optimization algorithm. To inform the gradients of its objective function, I will not use the model's output. However, the prediction models internal structure itself to optimally exploit the "hidden" information about the interactions with other pedestrians and the robot that are stored in the latent representations of the deeply learned model. Thereby the main objective of the algorithm is not to reach the goal in the smallest possible robot travel-time, but to disturb the surrounding humans as less as feasible while still reaching the robot's goal. Additionally, a Hamilton-Jacobi reachability based constraint will be used for guaranteeing a safe interaction. The algorithm should be real-time feasible running with a frequency of $10 Hz$, similar to related work discussed in the chapter \ref{text:related} (e.g. \cite{Chen2017}).
