\chapter*{Preface}
This thesis was carried out at the \ac{ASL} at Stanford University. First of all, I want to thank Prof. Marco Pavone for hosting me in his laboratory and supervising this work. Each discussion was insightful and helped shaping this project. I also want to thank Prof. Emilio Frazzoli for co-supervising this thesis and advising me during my studies at \ac{ETH}.
\newline
Working at \ac{ASL} was a very enriching experience and I want to thank all the members of the lab for welcoming me in this unique environment. I am deeply grateful to Karen Leung and Boris Ivanovic for their constant advice, for sharing their knowledge and experience about optimization, trajectory optimization and other related subjects.
\newline
When I arrived at Stanford, the field of (trajectory) optimization was obscure to me. I am very thankful for the help and for the fruitful discussions with several lab members as well as lab visitors; Tim Salzmann for the many discussions about machine learning, Riccardo Bonalli and Matteo Zallio for their advice and the insightful discussions over lunch, Robert Dyro for his support in efficient PyTorch programming, Jan Carius for his introduction to trajectory optimization, Jan Schiliger for his knowledge about control theory,  Thomas Lew and Layla Martin for their advice in (general) optimization techniques and Lina Fang for proof-reading my work. I will always be grateful to my family which has always been supportive throughout my life, encouraged me during my studies and greatly supported me throughout my thesis in this difficult times. Finally, I am grateful to the \ac{ETH} Excellence Scholarship and Opportunities Program as well as the Werner Siemens Foundation for financially supporting my stay at Stanford.

\cleardoublepage